Some examples in extended logic counting classes

### 2016-12-29 LICS/counting.tex
Arreglo apendice.

Then, what functional complexity classes can be captured with fragments of $\qso$?
For answering this question, it is reasonable to start with $\shp$ who is one of the most well-known functional complexity class~\cite{arora2009computational}. 
Since $\shp$ has a strong similarity with $\np$, one could expect a ``Fagin-like'' Theorem~\cite{fagin1974generalized} for this class. 
Actually, in~\cite{SalujaST95} it was shown that the class $\sfo$ captures $\shp$.
In our setting, the class $\sfo$ is contained in $\eqso(\fo)$ which also captures $\shp$ as expected. 
\begin{proposition} \label{prop:capture-shP}
	$\eqso(\fo)$ captures $\shp$ over ordered structures.
\end{proposition}
Recall that (see Section~\ref{sec:previous}) that every function class $\# \LL$ is contained in $\eqso(\LL)$. Thus, by~\cite{SalujaST95} it directly follows that every $\shp$-function can be defined in $\eqso(\LL)$. The other direction is also direct and follows by showing that $\shp$ is closed under first- and second-order sum.
Moreover, these closure properties also hold if we change the boolean core $\fo$ to any polynomial-time logic like $\lfp$. This gives the following corollary.
\begin{corollary} \label{eqso-lfp-cap}
	$\eqso(\lfp)$ captures $\shp$ over ordered structures.
\end{corollary}
By following the same approach as~\cite{SalujaST95}, Compton et al.~\cite{ComptonG96} show that $\seso$ captures $\spp$. As one could expect, if we parametrize $\eqso$ with $\eso$, we can also capture~$\spp$.
\begin{proposition} \label{eqso-eso-cap}
	$\eqso(\eso)$ captures $\spp$ over ordered structures.
\end{proposition}
The previous result connecting fragments from $\# \LL$ with $\shp$ and $\spp$ is the further that we can go by using SO-formulas with free variables. 
So, what happen with other counting complexity classes like $\fp$ and $\fpspace$?
By trying to search for an $\# \LL$ class for capturing $\fp$ one can easily see that this is pointless.
Indeed, by using free SO-variables one will easily capture $\shp$-hard functions (see Section~\ref{sec:syntactic}) and by only using free FO-variables one cannot capture exponential functions definable in $\fp$ (e.g. $2^{n}$).
Therefore, it is impossible to capture classes like $\fp$ or $\fpspace$ if we follow previous approaches like~\cite{SalujaST95}.

@@

Another natural function complexity class is $\fpspace$~\cite{Ladner89}, the functional version of $\pspace$. 
Of course, for capturing this class one needs a logical core powerful enough, like $\pfp$, for simulating the run of an $\pspace$-machine. 
On the other hand, an $\fpspace$ machine can generate outputs of exponentially many bits. 
For this reason, the second-order sum is not enough for the quantitative layer of a logic for $\fpspace$, that is, one needs to consider second-order products to produce big numbers. 
This leads us to the fragment $\qso(\pfp)$ which naturally captures $\fpspace$. 
\begin{theorem} \label{qso-pfp-cap}
	$\qso(\pfp)$ captures $\fpspace$ over ordered structures.
\end{theorem}
The proof of the previous theorem follows the same lines than the logical characterization of $\fp$: one shows that each function in $\qso(\pfp)$ can be computed in $\fpspace$ and, conversely, the output of an $\fpspace$ can be reconstructed by using $\pfp$ and second-order quantitative quantifiers.

>>>>

Then, what functional complexity classes can be captured with fragments of $\qso$?
For answering this question, it is reasonable to start with $\shp$ who is one of the most well-known functional complexity class~\cite{arora2009computational}. 
Since $\shp$ has a strong similarity with $\np$, one could expect a ``Fagin-like'' Theorem~\cite{fagin1974generalized} for this class. 
Actually, in~\cite{SalujaST95} it was shown that the class $\sfo$ captures $\shp$.
In our setting, the class $\sfo$ is contained in $\eqso(\fo)$ which also captures $\shp$ as expected. 
\begin{proposition} \label{prop:capture-shP}
	$\eqso(\fo)$ captures $\shp$ over ordered structures.
\end{proposition}
Recall that (see Section~\ref{sec:previous}) that every function class $\# \LL$ is contained in $\eqso(\LL)$. Thus, by~\cite{SalujaST95} it directly follows that every $\shp$-function can be defined in $\eqso(\LL)$. The other direction is also direct and follows by showing that $\shp$ is closed under first- and second-order sum.
Moreover, these closure properties also hold if we change the boolean core $\fo$ to any polynomial-time logic like $\lfp$. This gives the following corollary.
\begin{corollary} \label{prop:capture-shP-LFP}
	$\eqso(\lfp)$ captures $\shp$ over ordered structures.
\end{corollary}
By following the same approach as~\cite{SalujaST95}, Compton et al.~\cite{ComptonG96} show that $\seso$ captures $\spp$. As one could expect, if we parametrize $\eqso$ with $\eso$, we can also capture~$\spp$.
\begin{proposition} \label{prop:capture-spanP}
	$\eqso(\eso)$ captures $\spp$ over ordered structures.
\end{proposition}
The previous result connecting fragments from $\# \LL$ with $\shp$ and $\spp$ is the further that we can go by using SO-formulas with free variables. 
So, what happen with other counting complexity classes like $\fp$ and $\fpspace$?
By trying to search for an $\# \LL$ class for capturing $\fp$ one can easily see that this is pointless.
Indeed, by using free SO-variables one will easily capture $\shp$-hard functions (see Section~\ref{sec:syntactic}) and by only using free FO-variables one cannot capture exponential functions definable in $\fp$ (e.g. $2^{n}$).
Therefore, it is impossible to capture classes like $\fp$ or $\fpspace$ if we follow previous approaches like~\cite{SalujaST95}.

@@

Another natural function complexity class is $\fpspace$~\cite{Ladner89}, the functional version of $\pspace$. 
Of course, for capturing this class one needs a logical core powerful enough, like $\pfp$, for simulating the run of an $\pspace$-machine. 
On the other hand, an $\fpspace$ machine can generate outputs of exponentially many bits. 
For this reason, the second-order sum is not enough for the quantitative layer of a logic for $\fpspace$, that is, one needs to consider second-order products to produce big numbers. 
This leads us to the fragment $\qso(\pfp)$ which naturally captures $\fpspace$. 
\begin{theorem} \label{theo:capture-fpspace}
	$\qso(\pfp)$ captures $\fpspace$ over ordered structures.
\end{theorem}
The proof of the previous theorem follows the same lines than the logical characterization of $\fp$: one shows that each function in $\qso(\pfp)$ can be computed in $\fpspace$ and, conversely, the output of an $\fpspace$ can be reconstructed by using $\pfp$ and second-order quantitative quantifiers.
